Finite Complete Suites for CSP Refinement Testing

### 2018-11-26 scp-journal/article/complexity.tex
Considered message of 25.11

% -----------------------------------------------------------------------------
\paragraph{Test Cases of $\pmax$} The test cases $U_F(j)$ generated from $\pmax$ can
never enter branch (\ref{eq:ufa}), because $\pmax/s$ has initials $\Sigma$
for all traces $s\in\trc(\pmax)$ according to (\ref{eq:pmaxa}). Moreover,
they can never enter branch (\ref{eq:ufb}), because $\minhits(\pmax/s)$ is
never empty according to (\ref{eq:pmaxd}). Finally, the minimal hitting sets
used to probe the SUT at the end of a non-blocking test execution are always
the hitting sets of ${\cal C}$ according to (\ref{eq:pmaxe}). This results in
the following test case structure.
\begin{eqnarray*}
\label{eq:UFPpmax}
U_F(j) & = & U_F(j,0,\ii n)
\\
U_F(j,k,n) & = &   (k < j) \& \big(e:\Sigma   \then U_F(j,k+1,t(n,e) \big)
\\ & & \extchoice
\\ & & (k = j) \& \big( \Intchoice_{H\in\minhits({\cal C})} (e:H   \then \epass \then\Stop   )  \big)
\end{eqnarray*}

% -----------------------------------------------------------------------------
\paragraph{Maximal Number of Test Executions for $\pmax$}
When considering the number of test executions to be performed by $U_F(j)$
against some SUT $Q$ for all $j=0,\dots,pq-1$, the maximal number of test
executions is only reached if $Q$ is a correct refinement of the reference
process or if it fails in the very last execution of the very last $U_F(j)$
executed against $Q$. If an erroneous behaviour of $Q$ is revealed before
this very last execution, the test experiment is stopped (and $Q$ can be
fixed before executing the suite again). Therefore, we consider only correct
SUTs $Q$ when determining the maximal number of executions that can be
executed.

For the  $U_F(j)$ generated from $\pmax$, the number of test executions
 to be performed is maximal if $\pmax\lessdet_F Q$ and
$\trc(Q) = \Sigma^*$. In such a situation, no test execution blocks early, because
$Q/s$ can always engage into some $e\in\Sigma$   while $\#s<j$ and will never block
in the last step when $\#s = j$ and a hitting set $H\in\minhits({\cal C})$ is
offered by the test case. The resulting number of executions is
%
\begin{equation}
\label{eq:maxexec}
n^{j}\cdot \binom{n}{\lfloor\frac{n}{2}\rfloor},
\end{equation}
%
because all traces up to length $j$ can be executed with $U_F(j)$ entering branch
(\ref{eq:ufc}), and each of these traces is followed by one event
from each of the hitting sets of ${\cal C}$.

The number of executions specified in (\ref{eq:maxexec}) is indeed maximal for
all reference processes $P$ fulfilling $\minhits(P/s)\neq\varnothing$ for all
traces $s$: All these processes can never enter branch (\ref{eq:ufb}), and,
if an execution with an SUT entered branch (\ref{eq:ufa}), this would only lead to early termination of the whole test suite, because a failure has been detected.
As a consequence, $\card{\Sigma}^{j}$ is the maximal number of traces to be executed
up to length $j$, and from Theorem~\ref{th:sperner} we know that the number
$\binom{n}{\lfloor\frac{n}{2}\rfloor}$ of hitting sets to be tested at the end of each trace of length $j$ is already maximal.

Summing up  formula
(\ref{eq:maxexec}) over all test cases $U_F(0),\dots,U_F(pq-1)$ to be executed and applying the formula for the sum of a geometric progression,  this results in
%
\begin{equation}\label{eq:pmaxcomplexity}
\sum_{j=0}^{pq-1} n^{j}\cdot \binom{n}{\lfloor\frac{n}{2}\rfloor}  =
\binom{n}{\lfloor\frac{n}{2}\rfloor}\cdot\frac{1-n^{pq}}{1-n}\quad\text{with $n=\card{\Sigma}$}
\end{equation}
%
as the maximal number of test executions to be performed when testing an error-free
SUT $Q$ against reference process $\pmax$. This maximum is met when $\trc(Q) = \Sigma^*$.
If one is only interested in the order of magnitude, the maximal nmber of executions
may be re-written as
%
\begin{equation}
\label{eq:maxO}
O\big(\binom{n}{\lfloor\frac{n}{2}\rfloor}\cdot n^{pq-1}\big)\quad\text{with $n=\card{\Sigma}$}.
\end{equation}

% -----------------------------------------------------------------------------
\subsubsection*{Considering Empty Collections of Minimal Hitting Sets}

The argument so far has shown that reference process $\pmax$ requires the most test
executions among all CSP processes $P$ whose collections of minimal hitting sets
$\minhits(P/s)$ are never empty for any trace $s$, so that the branch
(\ref{eq:ufb}) of the related test cases $U_F(j)$ can never be entered.
Lemma~\ref{lemma:failureshittingsets} implies that an empty collection
$\minhits(P/s)$
is equivalent to $(s,A)\in\fails(P)$ for all $A\subseteq\Sigma$. This is equivalent to
$(s,\Sigma)\in\fails(P)$ or
$\maxrefs(P/s) = \{ \Sigma \}$ or $\minaccs(P/s) = \{ \varnothing\}$.

It remains to
consider the question whether reference processes $Z$ possessing failures $(s,\Sigma)$
may require more test executions for their associated tests $U_F(j)$ than the bound
given for $\pmax$ in (\ref{eq:pmaxcomplexity}), because process states $Z/s$ with
$\minhits(Z/s)=\varnothing$ allow for test executions entering branch (\ref{eq:ufb}).
To this end, consider a test case $U_F(j)$ constructed from such a process $Z$. Every trace
$s\in\trc(Z)$ with $\#s<j$ ending in a process state $Z/s$ with $\minhits(Z/s)=\varnothing$ allows for

>>>>

% -----------------------------------------------------------------------------
\paragraph{Test Cases of $\pmax$} The test cases $U_F(j)$ generated from $\pmax$ can
never enter branch (\ref{eq:ufa}), because $\pmax/s$ has initials $\Sigma$
for all traces $s\in\trc(\pmax)$ according to (\ref{eq:pmaxa}). Moreover,
they can never enter branch (\ref{eq:ufb}), because $\minhits(\pmax/s)$ is
never empty according to (\ref{eq:pmaxd}). Finally, the minimal hitting sets
used to probe the SUT at the end of a non-blocking test execution are always
the hitting sets of ${\cal C}$ according to (\ref{eq:pmaxe}). This results in
the following test case structure.
\begin{eqnarray*}
\label{eq:UFPpmax}
U_F(j) & = & U_F(j,0,\ii n)
\\
U_F(j,k,n) & = &   (k < j) \& \big(e:\Sigma   \then U_F(j,k+1,t(n,e) \big)
\\ & & \extchoice
\\ & & (k = j) \& \big( \Intchoice_{H\in\minhits({\cal C})} (e:H   \then \epass \then\Stop   )  \big)
\end{eqnarray*}
This means that the branches of $U_F(j)$ that can lead to an early
termination are not feasible. All tests deadlock, or run to the end of a
trace of size $j$ and then present the choice of event of a minimal hitting
set.

% -----------------------------------------------------------------------------
\paragraph{Maximal Number of Test Executions for $\pmax$}
When considering the number of test executions to be performed using $U_F(j)$
derived from $\pmax$ against some SUT $Q$ for all $j = 0,\dots,pq-1$, and
considering that we need to cover all the possible behaviours of $Q$, the
maximal number of test executions is only reached if (a)~$Q$ is a correct
refinement of $\pmax$ (or more generally, of the reference process) and
(b)~its traces are $\Sigma^*$.
%or if it fails in the very last execution of the very last $U_F(j)$
%executed against $Q$. If an erroneous behaviour of $Q$ is revealed before
%this very last execution, the test experiment is stopped (and $Q$ can be
%fixed before executing the suite again). Therefore, we consider only correct
%SUTs $Q$ when determining the maximal number of executions that can be
%executed.
%
%For the  $U_F(j)$ generated from $\pmax$, the number of test executions
% to be performed is maximal if $\pmax\lessdet_F Q$ and
%$\trc(Q) = \Sigma^*$.
In such a situation, no test execution blocks early, because $Q/s$ can always
engage into some $e\in\Sigma$  while $\#s<j$, and never blocks in the last
step when $\#s = j$ and a hitting set $H\in\minhits({\cal C})$ is offered by
the test case. The resulting number of executions in this case is
%
\begin{equation}
\label{eq:maxexec}
n^{j}\cdot \binom{n}{\lfloor\frac{n}{2}\rfloor},
\end{equation}
%
because \pagebreak all traces up to length $j$ can be executed with $U_F(j)$
entering branch (\ref{eq:ufc}), and each of these traces is followed by one
event from each of the hitting sets of ${\cal C}$ since $Q$ is correct.  %(b) leads to the maximal number of
%executions possible, because it is responsible for the factor *n^{(pq-1)} in
%the upper bound (47), (48) [n = |Sigma|]. Considering a reference process
%with fewer traces or an implementation Q with fewer traces would
%significantly reduce the value of this factor, while just adding one
%additional execution possibility for branch (20), after which the test suite
%is stopped due to failure.

The number of executions in (\ref{eq:maxexec}) is indeed maximal for all
reference processes $P$ fulfilling $\minhits(P/s)\neq\varnothing$ for all
traces $s$. All these processes can never enter branch (\ref{eq:ufb}), and,
if an execution with an SUT entered branch (\ref{eq:ufa}), this would only
lead to early termination of the whole test suite, because a failure has been
detected. As a consequence, $\card{\Sigma}^{j}$ is the maximal number of
traces to be executed up to length $j$, and from Theorem~\ref{th:sperner} we
know that the number $\binom{n}{\lfloor\frac{n}{2}\rfloor}$ of hitting sets
to be tested at the end of each trace of length $j$ is already maximal.

Summing up  formula (\ref{eq:maxexec}) over all test cases
$U_F(0),\dots,U_F(pq-1)$ to be executed and applying the formula for the sum
of a geometric progression,  this results int
%
\begin{equation}\label{eq:pmaxcomplexity}
\sum_{j=0}^{pq-1} n^{j}\cdot \binom{n}{\lfloor\frac{n}{2}\rfloor}  =
\binom{n}{\lfloor\frac{n}{2}\rfloor}\cdot\frac{1-n^{pq}}{1-n}\quad\text{with $n=\card{\Sigma}$}
\end{equation}
%
as the maximal number of test executions to be performed when testing an
error-free SUT $Q$ with $\trc(Q) = \Sigma^*$ against the reference process
$\pmax$. If we are interested only in the order of magnitude, the maximal
number of executions may be given as
%
\begin{equation}
\label{eq:maxO}
O\big(\binom{n}{\lfloor\frac{n}{2}\rfloor}\cdot n^{pq-1}\big)\quad\text{with $n=\card{\Sigma}$}.
\end{equation}

% -----------------------------------------------------------------------------
\paragraph{Considering Empty Collections of Minimal Hitting Sets}
The argument so far has shown that the tests derived from the reference
process $\pmax$ require the most test executions when compared to tests
derived from any other CSP process $P$ whose collections of minimal hitting
sets $\minhits(P/s)$ are never empty for any trace $s$, so that the branch
(\ref{eq:ufb}) of the related test cases $U_F(j)$ can never be entered.
Lemma~\ref{lemma:failureshittingsets} implies that an empty collection
$\minhits(P/s)$ is equivalent to $(s,A)\in\fails(P)$ for all
$A\subseteq\Sigma$. This is equivalent to $\maxrefs(P/s) = \{ \Sigma \}$ or
$\minaccs(P/s) = \{ \varnothing\}$.

It remains to consider whether reference processes $Z$ possessing failures
$(s,\Sigma)$ may require more test executions for their associated tests
$U_F(j)$ than the bound given for $\pmax$ in (\ref{eq:pmaxcomplexity}),
because process states $Z/s$ with $\minhits(Z/s)=\varnothing$ allow for test
executions entering branch (\ref{eq:ufb}). To this end, consider a test case
$U_F(j)$ constructed from such a process $Z$. Every trace $s\in\trc(Z)$ with
$\#s<j$ ending in a process state $Z/s$ with $\minhits(Z/s)=\varnothing$
allows for
